---
title       : "Surviving the Titantic"
subtitle    : "Would you survive or perish on the Titantic"
author      : Barrie D. Fitzgerald
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
github:
  user: barriefitzgerald
  repo: titantic_slidify
---

## Data Processing and Clean Up
```{r, echo=FALSE, warning=FALSE, message=FALSE}
setwd("C:/Users/bdfitzgerald/Desktop/Data Science Specialist/Developing Data Products")
library(caret)
library(rpart)
library(rattle)
library(rpart.plot)
library(RColorBrewer)
library(lsr)
```

-Data was obtained from kaggle.com's Titantic: Machine Learning from Disaster  
-Training data was loaded and partitioned into 60% for training and 40% crossvalid data set  
-Data clean up began
```{r,echo=TRUE, results='hide'}
set.seed(2015)
train <- read.csv("C:/Users/bdfitzgerald/Desktop/Data Science Specialist/Developing Data Products/train.csv")
intrain <- createDataPartition(train$Survived, p = 0.60, list = FALSE)
train <- train[intrain,]; crossvalid <- train[-intrain,]
```
```{r, echo=TRUE, results='asis'}
colnames(train)
```

--- .class #id 

## Survival and Perish Rates

```{r, echo=FALSE, results='hide'}
## Functions to clean data
cleanup <- function(x) {
        data <- x
        ## age bracket function
        age.bracket <- function(x) {
                train <- x
                ## separating out the age data
                age <- as.data.frame(train[, 6])
                colnames(age) <- "ageraw"
                ## Age Groups
                age$age <- age$ageraw
                age$age[is.na(age$ageraw) == TRUE] <- 999999        
                age$agebrack <- "Teen"
                age$agebrack[age$age < 11] <- "Child"
                age$agebrack[age$age >= 20] <- "Adult"
                age$agebrack[age$age >= 50] <- "Older Adult"
                age$agebrack[age$age == 999999] <- "Unknown"
                ## removing old age data
                age <- age[, -1]
                ## remaining data...nothing is done to it
                data <- train[, -6]
                ## combining the data back together
                train <- cbind(data, age)
        }
        ## applying the age bracket function
        data <- age.bracket(data)
        ## Personal Title Function
        titles <- function(x) {
                train <- x
                ## separating the names out
                titles <- as.data.frame(train$Name)
                colnames(titles) <- "names"
                titles$names <- as.character(titles$names)
                ## spliting names of indiviuals
                ## separate function
                separate <- function(x) {strsplit(x, split='[,.]')[[1]][2]}
                titles$titles <- sapply(titles$names,
                                        FUN = separate)
                titles$titles <- sub(" ", "", titles$titles)
                ## adjusting the names of the individuals
                titles$titles[titles$titles %in% c('Mme', 'Mlle')] <- 'Mlle'
                titles$titles[titles$titles %in% c('Capt', 'Don', 'Major', 'Sir')] <- 'Sir'
                titles$titles[titles$titles %in% c('Dona', 'Lady', 'the Countess', 'Jonkheer')] <- 'Lady'
                titles$titles[titles$titles %in% c('Mrs', 'Ms')] <- 'Mrs.Ms'
                titles$titles <- factor(titles$titles)
                ## remaining data...nothing is done with it
                data <- train[,-4]
                ## combining the data back together
                train <- cbind(data, titles)
        }
        ## applying the titles function
        data <- titles(data)
        ## Age adjustment function based on titles
        age.cleanup <- function(x) {
                train2 <- x
                data <- train2[, c(11, 14, 12)]
                ## adjusting ages based on titles
                data$agebrack.new <- data$agebrack
                data$agebrack.new[train2$age == 999999 &
                                          train2$titles == "Master"] <- "Child"
                data$agebrack.new[train2$age == 999999 &
                                          train2$titles == "Mr"] <- "Adult"
                data$agebrack.new[train2$age == 999999 &
                                          train2$titles == "Mrs"] <- "Adult"
                data$agebrack.new[train2$age == 999999 &
                                          train2$titles == "Miss"] <- "Adult"
                data$agebrack.new[train2$age == 999999 &
                                          train2$titles == "Ms"] <- "Adult"
                data$agebrack.new[train2$age == 999999 &
                                          train2$titles == "Mrs.Ms"] <- "Adult"
                ## keeping only the new age brackets
                data <- as.data.frame(data[, 4])
                colnames(data) <- "agebrack"
                ## remaining data...nothing done to it
                train2 <- train2[, -12]
                ## combine the data back together
                train2 <- cbind(train2, data)
        }
        ## applying the age bracket clean up function
        data <- age.cleanup(data)
}
train <- cleanup(train)
```
-Survival Rates by Sex
```{r, echo=FALSE}
knitr::kable(round(prop.table(table(train$Survived, train$Sex), 2),3))

```
-Survival Rates by Age Bracket
```{r, echo=FALSE, fig.align='center'}
knitr::kable(round(prop.table(table(train$Survived, train$agebrack), 2),3))

```
-Survival Rates by Ticket Class
```{r, echo=FALSE}
knitr::kable(round(prop.table(table(train$Survived, train$Pclass), 2),3))

```

--- .class #id 

## Model Algorithm
-A simple basic decision tree was generated using the Age Brackets, Sex, and Ticket Class
```{r, echo=FALSE, fig.align='center'}
## building the model
model <- rpart(formula = Survived ~ Sex + agebrack + Pclass, 
               data = train, control=rpart.control(minsplit=2, cp=0), 
               method = "class")
## plotting the modelt
fancyRpartPlot(model)
```

--- .class #id 

## Data Product

Using the algorithm, a data product was produced so an individual could find out if he or she would survive or perish if they were on the boat.   

http://barriefitzgerald.shinyapps.io/titantic_survival_prediction/


